---
title: "moma+tata data"
output: html_document
---

#  Packages and Libraries 
```{r}
install.packages("mapdata")
```
```{r}
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)

```

```{r}
library(dplyr)
library(stringr)
library(readr)
```

```{r}
library(RColorBrewer)
no_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )
```

#  MOMA 
##  Graph 1: Gender of Artists
 (decade)
```{r}
artists_moma <- read_csv("~/Desktop/data sets/MoMa collection/Artists.csv")
artists_moma = filter(artists_moma,!is.na(Gender))

artists_moma <- mutate(artists_moma,
                       Gender = ifelse(Gender %in% c("Female", "female"), "Female", "Male"))
 
  
artist_moma_count_histogram <- ggplot(data = artists_moma[artists_moma$BeginDate != 0,], aes(x = BeginDate))
```
### a) histogram 
```{r}
artist_moma_count_histogram + 
  geom_histogram(binwidth = 10, aes(fill = Gender)) +
  labs(title = "Who are the artists with works in MOMA?",
       subtitle = "Gender of artists through the decades",
       x = "Birth Dates of Artists",
       y = "Count") +
  theme_classic() +
  theme(plot.title= element_text(size = 16,
                                 family = "Palatino",
                                 hjust = 0.5),
        plot.subtitle = element_text(size = 14,
                                     family = "Palatino",
                                     hjust = 0.5),
        axis.title.x = element_text(size = 14,
                                    family = "Palatino"),
        axis.title.y = element_text(size = 14,
                                    family = "Palatino"),
        legend.title = element_text(family = "Palatino"),
        legend.text = element_text(family = "Palatino")) +
  scale_fill_manual(name = "Gender",
                     labels = c("Female", "Male"),
                     values = c("Female" = "gold2",
                                "Male" = "royalblue2" ))
```
### b) pie chart
```{r}

artist_moma_pie <- ggplot(artists_moma, aes(x = factor(1))) 

artist_moma_pie + 
  geom_bar(aes(x = 1, fill = Gender, width = 1)) +
  coord_polar("y", start = 0) +
  labs(title = "Who are the artists with works in MOMA?",
       subtitle = "Gender Composition of all artists") +
  theme_classic()+
  theme(plot.title= element_text(size = 16,
                                 family = "Palatino",
                                 hjust = 0.5),
        plot.subtitle = element_text(size = 14,
                                     family = "Palatino",
                                     hjust = 0.5),
        legend.title = element_text(family = "Palatino"),
        legend.text = element_text(family = "Palatino"),
        axis.text = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank()) +
  scale_fill_manual(name = "Gender",
                     labels = c("Female", "Male"),
                     values = c("Female" = "gold2",
                                "Male" = "royalblue2" ))
  
  # F-2360, 19%
  # M-10164, 81%
```

## Graph 2: Nationality and Gender

```{r}
artists_moma = filter(artists_moma,!is.na(Nationality))

nationality_moma <- artists_moma %>%                      #filter(artists_moma,!is.na(Nationality))%>%
                      group_by(Nationality)%>%
                      summarise(total = n())%>%
                      filter(total >= 150) %>%
                      ungroup()


nationality_moma_bar <- ggplot(data = artists_moma[artists_moma$Nationality %in% nationality_moma$Nationality,], aes(x = Nationality))

nationality_moma_bar +
  geom_bar(aes(fill = Gender), position = "fill") + 
  labs(title = "Who are the artists with works in MOMA?",
       subtitle = "Gender Makeup of Artists, Most Common Nationalities",
       x = "Nationality",
       y = "Proportion") +
  theme_classic() +
  theme(plot.title= element_text(size = 16,
                                 family = "Palatino",
                                 hjust = 0.5),
        plot.subtitle = element_text(size = 11,
                                     family = "Palatino",
                                     hjust = 0.5),
        axis.title.x = element_text(size = 14,
                                    family = "Palatino"),
        axis.title.y = element_text(size = 14,
                                    family = "Palatino"),
        axis.text.x = element_text(angle = 30,
                                   vjust = 1,
                                   hjust = 0.9),
        legend.title = element_text(family = "Palatino"),
        legend.text = element_text(family = "Palatino"),
        axis.ticks = element_blank(),
        axis.line = element_blank()) +
  scale_fill_manual(name = "Gender",
                     labels = c("Female", "Male"),
                     values = c("Female" = "gold2",
                                "Male" = "royalblue2" ))

```

## Graph 3: Is art around the world represented equally? Chlorapleth map 
```{r}
demonyms <- read_csv("~/Documents/GitHub/chunkedextractor/src/main/resources/edu/knowitall/chunkedextractor/demonyms.csv")
all_nationalities_moma <- artists_moma %>%
                      #filter(artists_moma,!is.na(Nationality))%>%
                      group_by(Nationality)%>%
                      summarise(total = n())

# rename demonyms col names, lost first row, but that country is 
# not in the moma list
names(demonyms) <- c("Nationality", "country_name")


all_nationalities_moma <- left_join(all_nationalities_moma, demonyms, by = "Nationality")

```

```{r}
world_map <- map_data('world') 

# rename USA and UK

all_nationalities_moma$country_name <- recode(all_nationalities_moma$country_name,
                                              'United States' = 'USA',
                                              'United Kingdom'= 'UK'
                                              )

#world_map_joined <- left_join(world_map, all_nationalities_moma, 
 #                             by = c('region' = 'country_name'))

world_map_joined <- left_join(all_nationalities_moma, world_map, 
                              by = c('country_name' = 'region'))
```

```{r}
cols_moma <- colorRampPalette(brewer.pal(11,'RdYlBu'))(length(world_map_joined))

moma_map <- ggplot()+
  geom_polygon(data = world_map, aes(long, lat, group = group),color = "grey28", size = 0.1) +
  geom_polygon(data = world_map_joined, aes(long, lat, group = group, fill = total),color = "grey28", size = 0.3) +  
  coord_fixed(1.1) +
  theme_classic() +
  no_axes +
  labs(title = "Who are the artists with works in MOMA?",
       subtitle = "Nationality of artists",
       legend = "Count") +
  theme(plot.title= element_text(size = 16,
                                 family = "Palatino",
                                 hjust = 0.5),
        plot.subtitle = element_text(size = 14,
                                     family = "Palatino",
                                     hjust = 0.5),
        legend.title = element_text(family = "Palatino"),
        legend.text = element_text(family = "Palatino"),
        axis.ticks = element_blank(),
        axis.line = element_blank()) +
  scale_fill_gradientn(colors = cols_moma, na.value = "black",trans = "log10")

moma_map
```
-- graph 3 v2 (DISCARDED, because found better palette in v1,) ----
------------ code kept just in case --------------
data_mid <- all_nationalities_moma[all_nationalities_moma$total < 1000,]
data_max <- all_nationalities_moma[all_nationalities_moma$total > 1000,]
data_min <- all_nationalities_moma[all_nationalities_moma$total <= 5,]

data_mid_joined <- left_join(world_map, data_mid, 
                              by = c('region' = 'country_name'))
data_max_joined <- left_join(data_max, world_map,
                              by = c('country_name'='region' ))
data_min_joined <- left_join(data_min, world_map,
                              by = c('country_name'='region'))

base <- ggplot(data_mid_joined, aes(long, lat, group = group)) + 
  geom_polygon(aes(fill=total),colour=NA, size = .3) + 
  scale_fill_gradient2(trans = "log10", high = rgb(102,102,255,maxColorValue
=255),low = rgb(255,102,128, maxColorValue =255),
  mid=rgb(178, 102, 225, maxColorValue =255),limits=c(5,1000)) + 
  coord_fixed()
  

ext <- base +
  geom_polygon(data = data_max_joined, fill = "blue") +
  geom_polygon(data = data_min_joined, fill = "red") 

ext
------------ code kept just in case --------------

# TATE
## Graph 1
```{r}
artist_tate <- read_csv("~/Desktop/data sets/tate collection ~2014/artist_data.csv")
artist_tate = filter(artist_tate,!is.na(gender))
```
### a) Histogram 
```{r}
  
artist_tate_count_histogram <- ggplot(data = artist_tate, aes(x = yearOfBirth))

artist_tate_count_histogram + 
  geom_histogram(binwidth = 10, aes(fill = gender)) +
  labs(title = "Who are the artists with works in Tate?",
       subtitle = "Gender of artists through the decades",
       x = "Birth Dates of Artists",
       y = "Count") +
  theme_classic() +
  theme(plot.title= element_text(size = 16,
                                 family = "Palatino",
                                 hjust = 0.5),
        plot.subtitle = element_text(size = 14,
                                     family = "Palatino",
                                     hjust = 0.5),
        axis.title.x = element_text(size = 14,
                                    family = "Palatino"),
        axis.title.y = element_text(size = 14,
                                    family = "Palatino"),
        legend.title = element_text(family = "Palatino"),
        legend.text = element_text(family = "Palatino")) +
  scale_fill_manual(name = "Gender",
                     labels = c("Female", "Male"),
                     values = c("Female" = "gold2",
                                "Male" = "royalblue2" ))

```
### b) Pie chart

```{r}
artist_tate_pie <- ggplot(artist_tate, aes(x = factor(1))) 

artist_tate_pie + 
  geom_bar(aes(x = 1, fill = gender, width = 1)) +
  coord_polar("y", start = 0) +
  labs(title = "Who are the artists with works in Tate?",
       subtitle = "Gender Composition of all artists") +
  theme_classic() +
  theme(plot.title= element_text(size = 16,
                                 family = "Palatino",
                                 hjust = 0.5),
        plot.subtitle = element_text(size = 14,
                                     family = "Palatino",
                                     hjust = 0.5),
        legend.title = element_text(family = "Palatino"),
        legend.text = element_text(family = "Palatino"),
        axis.text = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank()) +
  scale_fill_manual(name = "Gender",
                     labels = c("Female", "Male"),
                     values = c("Female" = "gold2",
                                "Male" = "royalblue2" ))

# F-521, 15%
# M-2895, 85%

```

## Graph 2 
```{r}

place_of_birth_split <- strsplit(artist_tate$placeOfBirth, ", ")
birth_country <- sapply(place_of_birth_split, tail, 1L)

unique_birthcountry <-data_frame(unique(birth_country))

artist_tate <- mutate(artist_tate, birth_country)

artist_tate = filter(artist_tate,!is.na(birth_country))
```

```{r}
#rename country names so it's consistent with universally accepted names, also bad coding in placeOfBirth column 
artist_tate$birth_country <- recode(artist_tate$birth_country,
                                    'Al-‘Iraq' = 'Iraq',
                                    "Al-Jaza'ir" = "Algeria",
                                    'Al-Lubnan' = 'Lebanon',
                                    'As-Sudan' = 'Sudan',  
                                    'Auteuil' = "France",
                                    'Beckington' = 'UK',
                                    'Brasil' = 'Brazil',
                                    'Bénin' = 'Benin',
                                    'België' = 'Belgium',
                                    'Bermondsey' = 'UK',
                                    'Bharat' = 'India',
                                    'Blackheath' = 'UK',
                                    'Bosna i Hercegovina' = 'Bosnia and Herzegovina',
                                    'Braintree' = 'UK',
                                    'Bristol' = 'UK',
                                    'Cameroun' = 'Cameroon',
                                    'Canterbury' = 'UK',
                                    'Ceská Republika' = 'Czech Republic',
                                    'Charlieu' = 'France',
                                    'Charlotte Amalie' = 'France',
                                    "Choson Minjujuui In'min Konghwaguk" = "North Korea",
                                    'Chung-hua Min-kuo' = 'China',
                                    'Danmark' = 'Denmark',
                                    'Deutschland'= 'Germany',
                                    'Douglas' = 'UK',
                                    'Edinburgh' = 'UK',
                                    'Eesti' = 'Estonia',
                                    'Egremont' = 'UK',
                                    'Éire' = 'UK',
                                    'Ellás' = 'Greece',
                                    'Epsom' = 'UK',
                                    'Hertfordshire' = 'UK',
                                    'Hrvatska' = 'Croatia',
                                    'Îran' = 'Iran',
                                    'Ísland' = 'Iceland',
                                    'Isle of Man' = 'UK',
                                    'Jugoslavija' = 'Yugoslavia',
                                    'Kensington' = 'UK',
                                    'Lao' = 'Laos',
                                    'Latvija' = 'Latvia',
                                    'Lietuva' = 'Lithuania',
                                    'Liverpool' = 'UK',
                                    'London' = 'UK',
                                    'Magyarország' = 'Hungary',
                                    'Makedonija' = 'Macedonia',
                                    'Melmerby' = 'UK',
                                    'México' = 'Mexico',
                                    'Misr' = 'Egypt',
                                    'Niederschlesien' = 'Germany',
                                    'Nihon' = 'Japan',
                                    'Norge' = 'Norway',
                                    'Panamá' = 'Panama',
                                    'Perth' = 'Australia',
                                    'Perú' = 'Peru',
                                    'Pilipinas' = 'Philipines',
                                    'Plymouth' = 'UK',
                                    'Prathet Thai' = 'Thailand',
                                    'Rochdale' = 'UK',
                                    'România' = 'Romania',
                                    'Saint Hélier' = 'Jersey',
                                    'Schlesien' = 'Poland',
                                    'Schweiz' = 'Switzerland',
                                    'Shqipëria' = 'Albania',
                                    'Österreich' = 'Austria',
                                    'Éire'= 'Ireland',
                                    'España'= 'Spain',
                                    'Italia' = 'Italy',
                                    'Polska' = 'Poland',
                                    'Nederland' = 'Netherlands',
                                    'Rossiya' = 'Russia',
                                    'Slovenija' = 'Slovenia',
                                    'Slovenská Republika' = 'Slovakia',
                                    'Solothurn' = 'Switzerland',
                                    'Staten Island' = 'USA',
                                    'United States' = 'USA',
                                    'Stockholm' = 'Sweden', 
                                    'Stoke on Trent' = 'UK',
                                    'Suomi' = 'Finland',
                                    'Suriyah' = 'Syria',
                                    'Sverige' = 'Sweden',
                                    "Taehan Min'guk" = "South Korea",
                                    'Türkiye' = 'Turkey',
                                    'Ukrayina' = 'Ukraine',
                                    'Viet Nam' = 'Vietnam',
                                    'Wimbledon' = 'UK',
                                    "Yisra'el" = 'Israel',
                                    'Zhonghua' = 'China'
                                              )

# nationality_tate: TOP nationalities of artists

nationality_tate <- artist_tate %>%
                      #filter(artists_moma,!is.na(birth_country))%>%
                      #select(birth_country, gender)
                      group_by(birth_country)%>%
                      summarise(total = n())%>%
                      filter(total >= 28) #taken top 13 nationalities



nationality_tate_bar <- ggplot(data = artist_tate[artist_tate$birth_country %in% nationality_tate$birth_country,], aes(x = birth_country))

nationality_tate_bar +
  geom_bar(aes(fill = gender), position = 'fill') +
  labs(title = "Who are the artists with works in Tate?",
       subtitle = "Gender Makeup of Artists, Most Common Nationalities",
       x = "Nationality",
       y = "Proportion") +
  theme_classic() +
  theme(plot.title= element_text(size = 16,
                                 family = "Palatino",
                                 hjust = 0.5),
        plot.subtitle = element_text(size = 11,
                                     family = "Palatino",
                                     hjust = 0.5),
        axis.title.x = element_text(size = 14,
                                    family = "Palatino"),
        axis.title.y = element_text(size = 14,
                                    family = "Palatino"),
        axis.text.x = element_text(angle = 30,
                                   vjust = 1,
                                   hjust = 0.9),
        legend.title = element_text(family = "Palatino"),
        legend.text = element_text(family = "Palatino"),
        axis.ticks = element_blank(),
        axis.line = element_blank()) +
  scale_fill_manual(name = "Gender",
                     labels = c("Female", "Male"),
                     values = c("Female" = "gold2",
                                "Male" = "royalblue2" ))

```

## Graph 3: Is art around the world represented equally? Chlorapleth map 
```{r}

all_nationalities_tate <- artist_tate %>%
                      group_by(birth_country)%>%
                      summarise(total = n())


world_map_joined_tate <- left_join(all_nationalities_tate, world_map,
                              by = c('birth_country' = 'region'))
```

```{r}
cols_tate <- colorRampPalette(brewer.pal(11,'RdYlBu'))(length(world_map_joined_tate))

tate_map <- ggplot()+
  geom_polygon(data = world_map, aes(long, lat, group = group),color = "grey28", size = 0.1) +
  geom_polygon(data = world_map_joined_tate, aes(long, lat, group = group, fill = total), color = "grey28") +  
  coord_fixed(1.1) +
  theme_classic() +
  no_axes +
  labs(title = "Who are the Artists with Works in Tate?",
       subtitle = "Nationality of Artists",
       legend = "Count") +
  theme(plot.title= element_text(size = 16,
                                 family = "Palatino",
                                 hjust = 0.5),
        plot.subtitle = element_text(size = 14,
                                     family = "Palatino",
                                     hjust = 0.5),
        legend.title = element_text(family = "Palatino"),
        legend.text = element_text(family = "Palatino"),
        axis.ticks = element_blank(),
        axis.line = element_blank()) +
  scale_fill_gradientn(colors = cols_tate, na.value = "black",trans = "log10")

tate_map
```

